Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-05-19 paper.tex
lots of changes and now it needs to be smaller

 How can an agent convert noisy, high-dimensional perceptual input
 to a symbolic, abstract object, such as a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting natural images, such as hand
 drawings, into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Figure~\ref{firstPageExamples}).
 %% \remark{The use of `graphics programs / visual programs' in the paper
 %%   title, title of this section, and the body of this section feels
 %%   too broad. `Graphics program' could conjur a lot of different ideas
 %%   (esp. 3D graphics); don't want to set the reader up to expect one
 %%   thing and then be disappointed that what you've done isn't
 %%   that. You bring up diagram-drawing later in the intro; I think it
 %%   should be made clear sooner (and certainly mentioned explicitly in
 %%   the abstract, when you get around to writing that).}

\newcommand{\exampleImageSize}{2cm}
\begin{figure}[h]
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \begin{minipage}[t]{0.3\linewidth}
\begin{verbatim}
for (i < 3)
 rectangle(3*i,-2*i + 4,
           3*i + 2,6)
 for (j < i + 1)
  circle(3*i + 1,-2*j + 5)
\end{verbatim}      
      \end{minipage}
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(bottom). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
  \end{figure}

High dimensional perceptual input may seem ill matched to the abstract
semantics of a programming language. But programs with constructs such as
recursion or iteration produce a simpler \emph{execution trace} of
primitive actions; for our domain, the primitive actions are drawing
commands. Our hypothesis is that the execution trace of the program is
better aligned with the perceptual input, and that the trace can act
as a kind of bridge between perception and programs. We test this
hypothesis by developing a model that learns to map from an image to
the execution trace of the graphics program that drew it.  With the
execution trace in hand, we can bring to bear techniques from the
program synthesis community to recover the latent graphics program.
This family of techniques, called \emph{constraint-based program synthesis}~\cite{solar2008program},
work by modeling a set of possible programs inside of a constraint solver,
such as a SAT or SMT solver~\cite{de2008z3}.
These techniques excel at uncovering high-level symbolic structure,
but are not well equipped to deal with real-valued perceptual inputs.

We develop a hybrid architecture for inferring graphics programs.  Our
approach uses a deep neural network infer an execution trace from an
image; this network recovers primitive drawing operations such as
lines, circles, or arrows, along with their parameters. For added
robustness, we use the deep network as a proposal distribution for a
stochastic search over execution traces.  Section~\ref{neuralNetworkSection} describes this first stage of the architecture where we infer drawing commands from images, and explains how we handle noisy hand drawings.
Finally, we use program synthesis techniques to recover the program from its
trace.  The program synthesizer discovers constructs such as loops and
geometric operations such as reflections and affine transformations.
% \remark{This paragraph is all about making things a bit more
Section~\ref{programSynthesisSection} describes how the architecture
synthesizes programs from execution traces,
and how those programs are used for  correcting errors made by the deep network, measuring similarity between hand drawings,
and extrapolating figures.

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given the features extracted by the
convnet, a multilayer perceptron then predicts a distribution over the
next drawing command to add to the trace.  We predict the drawing
command token-by-token, conditioning each token both on the image
features and on the previously generated tokens.  For example, the
network first decides to emit the \verb|circle| token conditioned on
the image features, then it emits the $x$ coordinate of the circle
conditioned on the image features and the \verb|circle| token, and
finally it predicts the $y$ coordinate of the circle conditioned on
the image features, the \verb|circle| token, and the $x$ coordinate.
See supplement for the full details of the architecture,
which we implemented in Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

The distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta [t_k | f_\theta(I,\text{render}(T)), \{t_j\}_{j = 1}^{k - 1}]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, and $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network). The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, and the subscripts
on $T$ index drawing commands within the trace, and the \verb|STOP|
token is emitted by the network to signal that the execution trace
explains the image.
%% ~\remark{Make explicit that a $T_n$ in Equation 2
%%   is a concise way of referring to a sequence of tokens from Equation
%%   1?}

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the execution trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}.}  \label{architecture}
\end{figure}
\begin{table}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes execution trace inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\caption{The deep network in (\ref{architecture}) predicts drawing commands, shown above.}
\label{drawingCommandTable}
\end{table}

We trained the network on $10^5$ scenes, which takes a little less than a day on a Nvidia TitanX GPU.

This network suffices to ``derender'' synthetic images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathcal{R}$ uses the distance between two images as a proxy for a
likelihood.

Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 8$
objects, it successfully parses scenes with many more objects than the
training data.

@@

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn reasonably carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid,
we made the drawings on graph paper.
\begin{figure}\centering
  \begin{minipage}{0.45\textwidth}
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}%
   \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
  \end{minipage}\hfill
  \begin{minipage}{0.45\textwidth}
  \includegraphics[width = 6cm]{figures/drawingAccuracy.png}
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1, 5, 10, 100 samples)?
  Distance to the ground truth trace is measured by the intersection over union of predicted vs. ground truth traces (sets of drawing commands).}\label{drawingIntersectionOverUnion}
    \end{minipage}
\end{figure}

@@

@@

\subsection{Correcting errors made by the neural network}\label{synthesizerHelpsParsing}
The program synthesizer can help correct errors from the execution trace proposal network by favoring execution traces which lead to more concise or general programs.
For example, one generally prefers figures with perfectly aligned objects over figures whose parts are slightly misaligned -- and precise alignment lends itself to short programs.
Similarly, figures often have repeated parts,
which the program synthesizer might be able to model as a loop or reflectional symmetry.
So, in considering several candidate traces proposed by the neural network,
we might prefer traces whose best programs have desirable features such being short or having iterated structures.

Concretely, we implemented the following scheme: for an image $I$, the neurally guided sampling scheme of section~\ref{neuralNetworkSection} samples a set of candidate traces, written $\mathcal{F}(I)$.
Instead of predicting the most likely trace in $\mathcal{F}(I)$ according to the neural network,
we can take into account the programs that best explain the traces. 
Writing $\hat{T}(I)$ for the trace the model predicts for image $I$,
\begin{equation}
\hat{T}(I) = \argmax_{T\in \mathcal{F}(I)} L_{\text{learned}}(I | \text{render}(T))\times \probability_\theta[T|I] \times\probability_{\beta} [ \text{program}(T)] 
\end{equation}
where $\probability_{\beta} [\cdot]$ is a prior probability
distribution over programs parameterized by $\beta$.
This is equivalent to doing
MAP inference in a generative model where the program is first drawn
from $\probability_{\beta} [\cdot]$, then the program is executed deterministically,
and then we observe a noisy version of the program's output, where $L_\text{learned}(I|\text{render}(\cdot))\times\probability_\theta[\cdot|I]$
is our observation model.

Given a corpus of graphics program synthesis problems with annotated ground truth traces (i.e. $(I,T)$ pairs),
we find a maximum likelihood estimate of $\beta$:
\begin{equation}
  \beta^* = \argmax_{\beta} \expect \left[ \log \frac{\probability_{\beta} [\text{program}(T)]\times L_{\text{learned}}(I|\text{render}(T))\times \probability_\theta[T|I]}{\sum_{T'\in \mathcal{F}(I)} \probability_{\beta} [\text{program}(T')]\times L_{\text{learned}}(I|\text{render}(T'))\times \probability_\theta[T'|I]} \right]
\end{equation}
where the expectation is taken both over the model predictions and the
$(I,T)$ pairs in the training corpus.  We define $\probability_{\beta}
[\cdot]$ to be a log linear distribution $\propto \exp
(\beta\cdot \phi(\text{program}))$, where $\phi(\cdot)$ is a feature
extractor for programs.  We extract a few basic features of a
program, such as its size and how many loops it has, and use these
features to help predict whether a trace is the correct explanation
for an image.

We synthesized programs for the top five traces
output by the deep network.  Learning this prior over programs can
help correct mistakes made by the neural network, and also
occasionally introduces mistakes of its own; see
Fig.~\ref{exampleOfProgramCorrectingMistake} for a representative
example of the kinds of corrections that it makes. On the whole
it modestly improves our Top-1 accuracy from 58\% to 60\%.  Recall that
from Fig.~\ref{drawingIntersectionOverUnion} that the best improvement
in accuracy we could possibly get is 63\% by looking at the top 5 traces.
\begin{wrapfigure}{R}{5.5cm}
  \includegraphics[width = 5cm]{figures/programSuccess7.png}
  \caption{Left: hand drawing. Center: interpretation favored by the deep network. Right: interpretation favored after learning a prior over programs. Our learned prior favors shorter, simpler programs, thus continuing the pattern of not having an arrow is preferred.}\label{exampleOfProgramCorrectingMistake}
  \end{wrapfigure}

@@

We measure the similarity between two drawings by extracting features
of the best programs that describe them.  Here the features we use are
just counts of the number of times that different components in the
DSL were used (Table~\ref{DSL}).  We project these features down to a
2-dimensional subspace using nonnegative matrix factorization
(NMF:~\cite{lee1999learning}); see Fig.\ref{NMF}.  One could use many
alternative similarity metrics between drawings which would capture pixel-level or object-level similarities while missing high-level geometric similarities.
For example, we can use our learned distance metric between execution traces, $L_{\text{learned}}(\cdot|\cdot)$.
Projecting these distances to a 2-dimensional subspace using multidimensional scaling (MDS:~\cite{cox2008multidimensional}) reveals similarities between the objects in the drawings,
while missing similarities at the level of the program.
\begin{figure}
  \begin{minipage}{0.48\textwidth}
    \includegraphics[width = \textwidth]{figures/NMF.png}
    \caption{NMF on features of the programs that were synthesized for each image. Horizontal component roughly corresponds to ``symmetry'' while vertical component roughly corresponds to ``loopyness'', with images on the diagonal having both of these.}    \label{NMF}
  \end{minipage}\hfill
  \begin{minipage}{0.48\textwidth}
    \includegraphics[width = \textwidth]{figures/imageSimilarity.png} 
    \caption{MDS on drawings using the learned distance metric, $L_{\text{learned}}(\cdot|\cdot)$. Drawings with similar looking parts in similar locations are clustered together.}
  \end{minipage}
\end{figure}

>>>>

 How can an agent convert noisy, high-dimensional perceptual input
 to a symbolic, abstract object, such as a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting natural images, such as hand
 drawings, into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Figure~\ref{firstPageExamples}a).
 %% \remark{The use of `graphics programs / visual programs' in the paper
 %%   title, title of this section, and the body of this section feels
 %%   too broad. `Graphics program' could conjur a lot of different ideas
 %%   (esp. 3D graphics); don't want to set the reader up to expect one
 %%   thing and then be disappointed that what you've done isn't
 %%   that. You bring up diagram-drawing later in the intro; I think it
 %%   should be made clear sooner (and certainly mentioned explicitly in
 %%   the abstract, when you get around to writing that).}

\newcommand{\exampleImageSize}{2cm}
\begin{figure}[h]
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \begin{minipage}[t]{0.3\linewidth}
\begin{Verbatim}
for (i < 3)
 rectangle(3*i,-2*i+4,
           3*i+2,6)
 for (j < i + 1)
  circle(3*i+1,-2*j+5)
\end{Verbatim}      
      \end{minipage}
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(bottom). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
\end{figure}

The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as boxes, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn; for example, in Figure 1(b), it means identifying that there is a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep neural network to infer a sequence of primitive shape-drawing commands that can generate an image similar to the observed image. We refer to this sequence as a \emph{program trace}, since it corresponds to the sequence of primitive commands that the desired program would have issued, but it lacks the high-level structure that determines how the program decided to issue those commands. For added robustness, we train the network to produce a proposal distribution of the most likely traces, and use stochastic search to find the trace that best matches the input. The network is trained from an automatically generated corpus of synthetic image generating programs.

The second step involves generating a high-level program capable of producing the program trace identified by the first phase. In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended. This paper argues that this stage is best achieved by \emph{constraint-based program synthesis} \cite{}. Without the need for training, the program synthesizer can search the space of possible programs for one capable of producing the desired trace under a quantitative objective that maximizes the likelihood according to a prior distribution that aims to capture the user’s preference.

%% The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as boxes, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn; for example, in Figure 1(b), it means identifying that there is a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

%% We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep neural network to infer a sequence of primitive shape-drawing commands that can generate an image similar to the observed image. We refer to this sequence as an \emph{execution trace}, since it corresponds to the sequence of primitive commands that the desired program would have issued, but it lacks the high-level structure that determines how the program decided to issue those commands. For added robustness, we train the network to produce a proposal distribution of the most likely traces, and use stochastic search to find the trace that best matches the input. The network is trained from an automatically generated corpus of synthetic image generating programs.

%% High dimensional perceptual input may seem ill matched to the abstract
%% semantics of a programming language. But programs with constructs such as
%% recursion or iteration produce a simpler \emph{execution trace} of
%% primitive actions; for our domain, the primitive actions are drawing
%% commands. Our hypothesis is that the execution trace of the program is
%% better aligned with the perceptual input, and that the trace can act
%% as a kind of bridge between perception and programs. We test this
%% hypothesis by developing a model that learns to map from an image to
%% the execution trace of the graphics program that drew it.  With the
%% execution trace in hand, we can bring to bear techniques from the
%% program synthesis community to recover the latent graphics program.
%% This family of techniques, called \emph{constraint-based program synthesis}~\cite{solar2008program},
%% work by modeling a set of possible programs inside of a constraint solver,
%% such as a SAT or SMT solver~\cite{de2008z3}.
%% These techniques excel at uncovering high-level symbolic structure,
%% but are not well equipped to deal with real-valued perceptual inputs.

%% We develop a hybrid architecture for inferring graphics programs.  Our
%% approach uses a deep neural network infer an execution trace from an
%% image; this network recovers primitive drawing operations such as
%% lines, circles, or arrows, along with their parameters. For added
%% robustness, we use the deep network as a proposal distribution for a
%% stochastic search over execution traces.  Section~\ref{neuralNetworkSection} describes this first stage of the architecture where we infer drawing commands from images, and explains how we handle noisy hand drawings.
%% Finally, we use program synthesis techniques to recover the program from its
%% trace.  The program synthesizer discovers constructs such as loops and
%% geometric operations such as reflections and affine transformations.
%% % \remark{This paragraph is all about making things a bit more
%% Section~\ref{programSynthesisSection} describes how the architecture
%% synthesizes programs from execution traces,
%% and how those programs are used for  correcting errors made by the deep network, measuring similarity between hand drawings,
%% and extrapolating figures.

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given the features extracted by the
convnet, a multilayer perceptron then predicts a distribution over the
next drawing command to add to the trace; see Table~\ref{drawingCommandTable}.  We predict the drawing
command token-by-token, conditioning each token both on the image
features and on the previously generated tokens.  For example, the
network first decides to emit the \verb|circle| token conditioned on
the image features, then it emits the $x$ coordinate of the circle
conditioned on the image features and the \verb|circle| token, and
finally it predicts the $y$ coordinate of the circle conditioned on
the image features, the \verb|circle| token, and the $x$ coordinate.
See supplement for the full details of the architecture,
which we implemented in Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

The distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta [t_k | f_\theta(I,\text{render}(T)), \{t_j\}_{j = 1}^{k - 1}]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, and $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network). The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, the subscripts
on $T$ index drawing commands within the trace, and the \verb|STOP|
token is emitted by the network to signal that the execution trace
explains the image.
%% ~\remark{Make explicit that a $T_n$ in Equation 2
%%   is a concise way of referring to a sequence of tokens from Equation
%%   1?}

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the execution trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}.}  \label{architecture}
\end{figure}
\begin{table}[h]
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes execution trace inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\caption{The deep network in (\ref{architecture}) predicts drawing commands, shown above.}
\label{drawingCommandTable}
\end{table}

We trained the network on $10^5$ scenes, which takes a little less than a day on an Nvidia TitanX GPU.

This network suffices to ``derender'' synthetic images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathcal{R}$ uses the distance between two images as a proxy for a
likelihood.
%
Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 8$
objects, it successfully parses scenes with many more objects than the
training data.

@@

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn reasonably carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid,
we made the drawings on graph paper.
\begin{figure}\centering
  \begin{minipage}{0.45\textwidth}
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}%
   \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
  \end{minipage}\hfill
  \begin{minipage}{0.45\textwidth}
  \includegraphics[width = 6cm]{figures/drawingAccuracy.png}
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1,5,10,100)?
  Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces (sets of drawing commands).}\label{drawingIntersectionOverUnion}
    \end{minipage}
\end{figure}

@@

\newcommand{\exampleProgramSize}{2.3cm} 
\begin{figure}
  \begin{tabular}{lllll}
  \includegraphics[width = \exampleProgramSize]{figures/expert-29.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-52.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-38.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-72.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-75.png}\\
    \begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize = \small]
for(i<3){
line(i,-1*i+6,
2*i+2,-1*i+6)
line(i,-2*i+4,
i,-1*i+6)
\end{Verbatim}
    \end{minipage}&
    \begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize = \small]
circle(4,10)
for(i<3){
circle(-3*i+7,5)
circle(-3*i+7,1)
line(-3*i+7,4,
-3*i+7,2,
arrow=True)
line(4,9,
-3*i+7,6,
arrow=True)
\end{Verbatim}
\end{minipage}&
\begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize=\small]
for(i<3)
for(j<3)
if(j>0)
line(-3*j+8,
-3*i+7,-3*j+9,
-3*i+7)
line(-3*i+7,-3*j+8,
-3*i+7,
-3*j+9)
circle(-3*j+7,-3*i+7)
\end{Verbatim}
\end{minipage}&
\begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize=\small]
reflect(y=8){
for(i<3){
if(i>0){
rectangle(3*i-1,
2,3*i,3)}
circle(3*i+1,
       3*i+1)
\end{Verbatim}
\end{minipage}&
 \begin{minipage}[t]{\exampleProgramSize}
    \begin{Verbatim}[fontsize = \small]
for(i<4){
line(-4*i+13,4,
-4*i+13,2,
arrow=True)
for(j<3){
if(j>0){
circle(-4*i+13,
4*j+-3)}
line(-4*j+10,5,
-4*j+12,5,
arrow=True)
\end{Verbatim}
  \end{minipage}
  \end{tabular}
  \caption{Example drawings (top row) and the programs we synthesize from their ground truth traces (bottom row). Notice the nested loops in the Ising model (middle), special case conditionals for the HMM (rightmost), combination of symmetry and iteration in middle left, and affine transformation in the leftmost figure.}
  \end{figure}

@@

\subsection{Correcting errors made by the neural network}\label{synthesizerHelpsParsing}
The program synthesizer can help correct errors from the execution
trace proposal network by favoring execution traces which lead to more
concise or general programs.  For example, one generally prefers
figures with perfectly aligned objects over figures whose parts are
slightly misaligned -- and precise alignment lends itself to short
programs.  Concretely, we estimated a prior over programs. Then,
given the few most likely traces output by the neurally guided sampler,
we reranked them according to the prior probability of their programs.
Our neurally guided sampler could only do better on
7 drawings by looking at the Top-100 sampled traces
(see Fig.~\ref{drawingIntersectionOverUnion}),
precluding a statistically significant analysis of how much
learning a prior over programs could help correct errors.
But,
learning this prior does sometimes
help correct mistakes made by the neural network, and also
occasionally introduces mistakes of its own; see
Fig.~\ref{exampleOfProgramCorrectingMistake} for a representative
example of the kinds of corrections that it makes.
See supplement for details.
\begin{wrapfigure}{R}{5.5cm}
  \includegraphics[width = 5cm]{figures/programSuccess7.png}
  \caption{Left: hand drawing. Center: interpretation favored by the deep network. Right: interpretation favored after learning a prior over programs. Our learned prior favors shorter, simpler programs, thus continuing the pattern of not having an arrow is preferred.}\label{exampleOfProgramCorrectingMistake}
  \end{wrapfigure}

%% This is equivalent to doing
%% MAP inference in a generative model where the program is first drawn
%% from $\probability_{\beta} [\cdot]$, then the program is executed deterministically,
%% and then we observe a noisy version of the program's output, where $L_\text{learned}(I|\text{render}(\cdot))\times\probability_\theta[\cdot|I]$
%% is our observation model.

%% Learning this prior over programs can
%%  On the whole
%% it modestly improves our Top-1 accuracy from 58\% to 61\%.  Recall that
%% from Fig.~\ref{drawingIntersectionOverUnion} that the best improvement
%% in accuracy we could possibly get is 65\% by looking at the top 10 traces.

@@

We measure the similarity between two drawings by extracting features
of the best programs that describe them. Our features are counts of the number of times that different components in the
DSL were used (Table~\ref{DSL}).  We project these features down to a
2-dimensional subspace using nonnegative matrix factorization
(NMF:~\cite{lee1999learning}); see Fig.\ref{NMF}.  One could use many
alternative similarity metrics between drawings which would capture pixel-level while missing high-level geometric similarities.
We used our learned distance metric between execution traces, $L_{\text{learned}}(\cdot|\cdot)$, and projected to a 2-dimensional subspace using multidimensional scaling (MDS:~\cite{cox2008multidimensional}). This reveals similarities between the objects in the drawings,
while missing similarities at the level of the program.
\begin{figure}
  \begin{minipage}{0.48\textwidth}
    \includegraphics[width = \textwidth]{figures/NMF.png}
    \caption{NMF on features of the programs that were synthesized for each image. Horizontal component roughly corresponds to ``symmetry'' while vertical component roughly corresponds to ``loopyness'', with images on the diagonal having both of these.}    \label{NMF}
  \end{minipage}\hfill
  \begin{minipage}{0.48\textwidth}
    \includegraphics[width = \textwidth]{figures/imageSimilarity.png} 
    \caption{MDS on drawings using the learned distance metric, $L_{\text{learned}}(\cdot|\cdot)$. Drawings with similar looking parts in similar locations are clustered together.}
  \end{minipage}
\end{figure}
